Bounds on Coding Theory\\from Algebraic Geometry

### 2021-05-01 draft.tex
2021.05.01 update

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

@@

\title[Draft: Codes from Algebraic Geometry]{Draft: Bounds on Coding Theory\\from Algebraic Geometry}
\author{Guilherme Zeus Dantas e Moura}
\date{2021-04-09} % <---- Change this date to the format YYYY-MM-DD

@@

\begin{prop} \label{prop:hamming}
    Hamming distance is a metric over $A^n$, i.e., the following holds for any $\mathbf{x}, \mathbf{y}, \mathbf{z} \in A^n$:
    \begin{itemize}[noitemsep]
        \item $\dist(\mathbf{x}, \mathbf{y}) = 0 \iff \mathbf{x} = \mathbf{y}$;
        \item $\dist(\mathbf{x}, \mathbf{y}) = \dist(\mathbf{x}, \mathbf{y})$;
        \item $\dist(\mathbf{x}, \mathbf{y}) \le \dist(\mathbf{x}, \mathbf{z}) + \dist(\mathbf{z}, \mathbf{y})$.
    \end{itemize}
\end{prop}

\begin{defn}[Parameters of a code]
    If $C$ is a linear code over $A$, we define \emph{dimension of $C$} as $k = \dimension_A(C)$ and \emph{minimum distance of $C$} as $d = \min\left\{\dist(\mathbf{x}, \mathbf{y}) \mid \mathbf{x}, \mathbf{y} \in C\right\}$.
	(\emph{If $C$ is a nonlinear code over an alphabet with size $q$, we can coherently define $k = \log_q{\left|C\right|}$.})
	The length $n$, dimension $k$ and minimum distance $d$ are the \emph{parameters} of $C$.
\end{defn}

@@

\section{Singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then 
	\begin{equation}
		k + d \le n + 1,
	\end{equation} 
	or equivalently, 
	\begin{equation}
		R + \delta \le 1 + 1/n.
	\end{equation}
\end{thm}

\begin{proof}
	We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code.\todo{Write proof}
\end{proof}

@@

The Reed--Solomon codes are very good codes in the sense that key have the largest possible sum $k + d$ for their length $n$.
However, Reed--Solomon codes are limited because their length is at most the alphabet size.
So, a question naturally arises: Given fixed $\mathbb{F}_q$, are there codes over $\mathbb{F}_q$ with arbitrarily large $n$ and $R + \delta = 1 + 1/n$?
If not, how large can $R$ and $\delta$ be when $n$ gets larger?
The well-known Gilbert--Varshamov bound shows that there are codes with 
\begin{equation}
	1 - R \approx H_q(\delta), \text{ as $n \to \infty$,}
\end{equation}
in which
\begin{equation}
	H_q(x) = x \log_q(q-1) - x\log_q(x) - (1-x)\log_q(x).
\end{equation}
The article \cite{TVZ82} uses Algebraic Geometry to show that there are codes that give better bounds.

@@

\begin{defn}[Algebraic curves]
	An affine algebraic plane curve $\mathfrak C_f$ is the zero set of a polynomial $f \in K[x, y]$.
	A projective algebraic plane curve $\mathfrak C_F$ is the zero set in a projective plane of a homogeneous polynomial $F \in K[X, Y, Z]$.
	Generally, algebraic curves are algebraic varieties of dimension $1$.
\end{defn}

@@

\begin{defn}[Rational points, \cite{Wal00}] %Not word-by-word
	Let $\mathfrak C$ be a projective curve defined by $F(X, Y, Z) = 0$, where $F \in K[X, Y, Z]$ is homogeneous polynomial.
	A point $(X_0 : Y_0 : Z_0) \in \mathbb{P}^2(L)$ is called a \emph{$L$-rational point} on $\mathfrak C$ if $F(X_0, Y_0, Z_0) = 0$.
	Additionally, if $(X_0 : Y_0 : Z_0)$ is also in $\mathbb{P}^2(K)$, we call it simply a \emph{rational point} on $\mathfrak C$.
\end{defn}

\begin{thm}[Bezout's theorem, \cite{Wal00}] %Not word-by-word
	If $F, G \in K[X, Y, Z]$ are homogeneous polynomials of degrees $d$ and $e$, respectively, then their curves $\mathfrak C_F$ and $\mathfrak C_G$ intersect in exactly $de$ $L$-rational points, counting multiplicity.
\end{thm}

\begin{defn}[Rational function]
	A rational function $f$ over a $K$ is the ratio $g/h$ of polynomials $g, h \in K[x, y]$.
	A homogeneous rational function $F$ over a $K$ is the ratio $G/H$ of homogeneous polynomials $G, H \in K[x, y]$.

	After cancelling common roots of $g, h$, the roots of $g$ are called \emph{zeros} of $f$ and the roots of $h$ are called the \emph{poles} of $f$.

	We say $f$ has order $n$ in $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.
\end{defn}

\begin{defn}[Divisor] 
	Let $\overline{\mathbb{F}_q}$ be the algebraic closure of $\mathbb{F}_q$.
	Let $\mathfrak C$ be a projective plane curve defined over $\mathbb{F}_q$.
	A \emph{divisor} $D$ on $\mathfrak C$ is a formal finite sum of the form $D = \sum a_P P$, where $P$ are points of $\mathfrak C$, $a_P$ are integers and $a_P = 0$ for all but finitely many points $P$.
	The \emph{degree} of $D$ is $\sum n_P$.
	The \emph{support} $\Supp D$ is the set $\{P \in X : a_P \neq 0\}$.
\end{defn}

@@

\begin{defn}[Algebraic geometric codes, \cite{LS87, Wal00}] %Copied from Wal00
	Let $\mathfrak C$ be an irreducible nonsingular%
	\todo{Provide an explanation for why the `irreducible nonsingular' is there, but not a thing that the reader should worry. I think the key is some sort of multiplicity.}
	projective plane curve, let $D$ be a divisor on $\mathfrak C$ and let $\mathcal P = \{P_1, P_2, \dots, P_n\}$ be a set of rational points on $\mathfrak C$.
	Assume $\mathcal P$ and $\Supp D$ have no points in common, thus, no $P_i$ can be a pole of any $f \in \mathcal L(D)$, and $f(P_i)$ is well-defined for any $f \in \mathcal L(D)$ and any $P_i \in \mathcal P$.
	Then, the \emph{algebraic geometric code} associated to $\mathfrak C, \mathcal P$ and $D$ is
	\begin{equation}
		C(\mathfrak C, \mathcal P, D) := \{(f(P_1), \dots, f(P_n)) \mid f \in L(D)\} \subset \mathbb{F}_q^n.
	\end{equation}
\end{defn}

>>>>

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem*{exmp*}{Example}

@@

\title[Codes from Algebraic Geometry]{Bounds on Coding Theory\\from Algebraic Geometry}
\author{Guilherme Zeus Dantas e Moura}
\date{2021-05-07} % <---- Change this date to the format YYYY-MM-DD

@@

\begin{prop} \label{prop:hamming}
    Hamming distance is a metric over $A^n$, i.e., the following holds for any $\mathbf{x}, \mathbf{y}, \mathbf{z} \in A^n$:
    \begin{itemize}[noitemsep]
        \item $\dist(\mathbf{x}, \mathbf{y}) = 0 \iff \mathbf{x} = \mathbf{y}$;
        \item $\dist(\mathbf{x}, \mathbf{y}) = \dist(\mathbf{y}, \mathbf{z})$;
        \item $\dist(\mathbf{x}, \mathbf{y}) \le \dist(\mathbf{x}, \mathbf{z}) + \dist(\mathbf{z}, \mathbf{y})$.
    \end{itemize}
\end{prop}

\begin{defn}[Parameters of a code]
	If $C$ is a code over $A$, we define \emph{dimension of $C$} as $k = \log_{|A|}|C|$ and \emph{minimum distance of $C$} as $d = \min\left\{\dist(\mathbf{x}, \mathbf{y}) \mid \mathbf{x} \neq \mathbf{y} \in C\right\}$.
	(\emph{If $C$ is a linear code, the definition above is equivalent to $k = \dimension_{A}(C)$.})
	The length $n$, dimension $k$ and minimum distance $d$ are the \emph{parameters} of $C$.
\end{defn}

@@

\section{Singleton bound and a promising example} \label{s:singleton}
\begin{thm}[Singleton Bound]\label{thm:singleton_bound}
If $C$ is a code with parameters $n, k, d$, then 
	\begin{equation} % IF SPACE NEEDED, TRANSFORM THESE EQUATIONS INTO INLINE
		k + d \le n + 1,
	\end{equation} 
	or equivalently, 
	\begin{equation}
		R + \delta \le 1 + 1/n.
	\end{equation}
\end{thm}

We will provide the proof for Theorem \ref{thm:singleton_bound} when $C$ is a linear code over a field $K$.

\begin{proof}
	Let $W := \{\mathbf{x} = (x_1, \dots, x_n) \mid x_{d} = \dots = x_n = 0\}$, which is a vector subspace of $K^n$ with dimension $d - 1$. Since $C$ is a vector space, $\vec 0 \in C$. For any non-zero vector $\vec v \in C$, $\dist(\vec 0, \vec v) \ge d$, thus $\vec v$ has at least $d$ non-zero entries, and therefore $\vec v \notin W$. Thus, $W \cap C = \{\vec 0\}$.

	Let $\vec w_1, \dots, \vec w_{d-1}$ and $\vec v_1, \dots, \vec v_k$ be a choice of basis for $W$ and $C$, respectively. Since $W \cap C = \{\vec 0\}$, the vectors $\vec w_1, \dots, \vec w_{d-1}, \vec v_1, \dots, \vec v_k$ are linearly independent. They all are vectors in $K^n$, therefore, $k + d - 1 \le n$.
\end{proof}

@@

The Reed--Solomon codes are very good codes in the sense that they have the largest possible sum $k + d$ for their length $n$.
However, Reed--Solomon codes are limited because their length is at most the alphabet size.
So, a question naturally arises: Given fixed $\mathbb{F}_q$, are there codes over $\mathbb{F}_q$ with arbitrarily large $n$ and $R + \delta = 1 + 1/n$?
If not, how large can $R$ and $\delta$ be when $n$ gets larger?
The well-known Gilbert--Varshamov bound shows that there are codes with 
\begin{equation}
	1 - R \approx H_q(\delta), \text{ as $n \to \infty$,}
\end{equation}
in which
\begin{equation}
	H_q(x) = x \log_q(q-1) - x\log_q(x) - (1-x)\log_q(x).
\end{equation}
The article \cite{TVZ82} uses Algebraic Geometry to show that there are codes that give better bounds.

@@

\begin{defn}[Algebraic plane curves]
	An affine algebraic plane curve $\mathfrak C_f$ is the zero set of a polynomial $f \in K[x, y]$.
	A projective algebraic plane curve $\mathfrak C_F$ is the zero set in a projective plane of a homogeneous polynomial $F \in K[X, Y, Z]$.\footnote{In more general terms, algebraic curves (as opposed to algebraic plane curves) are algebraic varieties of dimension $1$. For this paper, we will not worry about this; our discussion will be based on algebraic plane curves, but the results also follow in higher dimensions.}
\end{defn}

@@

\begin{defn}[Rational points, \cite{Wal00}] %Not word-by-word
	Let $\mathfrak C$ be a projective curve defined by $F(X, Y, Z) = 0$, where $F \in K[X, Y, Z]$ is a homogeneous polynomial.
	A point $(X_0 : Y_0 : Z_0) \in \mathbb{P}^2(L)$ is called a \emph{$L$-rational point} on $\mathfrak C$ if $F(X_0, Y_0, Z_0) = 0$.
	Additionally, if $(X_0 : Y_0 : Z_0)$ is also in $\mathbb{P}^2(K)$, we call it simply a \emph{rational point} on $\mathfrak C$.
\end{defn}

\begin{exmp*}
	Consider the curve $\mathfrak C$ defined by $X^2 + Y^2 + Z^2 = 0$ over $\mathbb{F}_7$.
	The point $(1 : 2 : 3)$ is a rational point on $\mathfrak C$.
	Note that the polynomial $x^2 - 3$ is irreducible over $\mathbb{F}_7$, but there it has a root $\alpha$ in its algebraic closure, $\overline{\mathbb{F}_7}$. The projective point $(1 :  \alpha : \alpha)$ is a $\overline{\mathbb{F}_7}$-rational point.
\end{exmp*}

\begin{thm}[Bezout's theorem, \cite{Wal00}] %Not word-by-word
	If $F, G \in K[X, Y, Z]$ are homogeneous polynomials of degrees $d$ and $e$, respectively, then their curves $\mathfrak C_F$ and $\mathfrak C_G$ intersect in exactly $d\cdot e$ $L$-rational points, counting multiplicity.
\end{thm}

\begin{defn}[Rational function]
	A rational function $f$ over $K$ is the ratio $g/h$ of polynomials $g, h \in K[x, y]$.
	A homogeneous rational function $F$ over a $K$ is the ratio $G/H$ of homogeneous polynomials $G, H \in K[x, y]$.

	After cancelling common roots of $g, h$, the roots of $g$ are called \emph{zeros} of $f$ and the roots of $h$ are called the \emph{poles} of $f$.  
	We say $f$ has order $n$ at $P$ if $P$ is a zero of multiplicity $n$; order $-n$ if $P$ is a pole with multiplicity $n$; and order $0$, otherwise.
\end{defn}

\begin{defn}[Divisor] 
%	Let $\overline{\mathbb{F}_q}$ be the algebraic closure of $\mathbb{F}_q$.
%	Let $\mathfrak C$ be a projective plane curve defined over $\mathbb{F}_q$.
	Let $\mathfrak C$ be a projective plane curve defined over $K$.
	A \emph{divisor} $D$ on $\mathfrak C$ is a formal finite sum of the form $D = \sum a_P P$, where $P$ varies over the $L$-rational points on $\mathfrak C$, $a_P$ is an integer and $a_P = 0$ for all but finitely many points $P$.
	The \emph{degree} of $D$ is $\sum a_P$.
	The \emph{support} $\Supp D$ is the set $\{P \in X : a_P \neq 0\}$.
\end{defn}

@@

\begin{defn}[Algebraic geometric codes, \cite{LS87, Wal00}] %Copied from Wal00
	Let $\mathfrak C$ be an irreducible nonsingular%
	\footnote{The adjectives make sure $\mathfrak C$ is well-behaved. The reader should not worry about them.}
%%	\todo{Provide an explanation for why the `irreducible nonsingular' is there, but not a thing that the reader should worry. I think the key is some sort of multiplicity.}
	projective plane curve, let $D$ be a divisor on $\mathfrak C$ and let $\mathcal P = \{P_1, P_2, \dots, P_n\}$ be a set of rational points on $\mathfrak C$.
	Assume $\mathcal P$ and $\Supp D$ have no points in common, thus, no $P_i$ can be a pole of any $f \in \mathcal L(D)$, and $f(P_i)$ is well-defined for any $f \in \mathcal L(D)$ and any $P_i \in \mathcal P$.
	Then, the \emph{algebraic geometric code} associated to $\mathfrak C, \mathcal P$ and $D$ is
	\begin{equation}
		C(\mathfrak C, \mathcal P, D) := \{(f(P_1), \dots, f(P_n)) \mid f \in L(D)\} \subset \mathbb{F}_q^n.
	\end{equation}
\end{defn}
